---
output: github_document
always_allow_html: yes
---
# Tidy eval and dplyr/tidyverse examples
#### A growing gallery of tidy NSE
#### Because of scoping, this will also include the dplyr "<operation>_at" functions

## Helpful links
The 2 iterations of Hadley's advanced R
<br>http://adv-r.had.co.nz/
<br>https://adv-r.hadley.nz/

The programming dplyr article
<br>https://dplyr.tidyverse.org/articles/programming.html
```{r boilerplate}
library(tidyverse)
library(ggvis)
```
### Compare multiple quosures
```{r Sample Data}
### These are not the same
quos(B, C, D) %>% 
  class()

list(quo(A), quo(B), quo(C)) %>% 
  class()

## Fails
# list(A, B, C)
```
### Create a tibble
#### 1. Single column, outside of function
```{r}
num_rows = 8

col_names = quo(A) %>% 
  quo_name()

tibble(!! col_names := rnorm(num_rows))
```
#### 2. Multiple columns 
Use map and quo_name
```{r}
col_names = quos(A, B, C, D) %>% 
  map(function(expr) quo_name(expr))

col_names %>% 
  map(function(col_name) tibble(!! col_name := rnorm(4))) %>% 
  bind_cols()
```
#### 2. Multiple columns without quo_name
```{r}
quos(A, B, C) %>% 
  map(function(col_name) tibble(!! col_name := rnorm(3))) %>% 
  bind_cols()
```
#### 4. In a function using ...
```{r}
create_tibble1 = function(col_names_f, num_rows_f = 100, ...){
  col_names_f = enquos(col_names_f, ...) %>% 
    map(function(expr) quo_name(expr))
  
  col_names_f %>%
    map(function(col_name) tibble(!! col_name := rnorm(num_rows_f))) %>% 
    bind_cols()
}

create_tibble1(col_names_f = B, C, D, E, num_rows_f = 6)
```
#### 5. In a function not using ...
```{r}
create_tibble2 = function(col_names_f, num_rows_f = 100){
  col_names_f = col_names_f %>% 
    map(function(expr) quo_name(expr))
  
  col_names_f %>%
    map(function(col_name) tibble(!! col_name := rnorm(num_rows_f))) %>% 
    bind_cols()
}

create_tibble2(col_names_f = quos(B, C, D, E), num_rows_f = 10)
```

### Group_by_at
```{r}
tibble(col1 = c("A", "B", "C") %>% 
         sample(size = 100, replace = TRUE)) %>% 
  group_by_at(names(.)) %>% 
  summarise(n = n())
```

### Graph examples
#### ggvis, which uses formulas
```{r}
set.seed(1)
df = tibble(a = 1:10,
            b = rnorm(10),
            c = rnorm(10))

# df %>% 
#   ggvis(~b, ~c) %>% layer_points()

ggvis_sample = function(data, col1, col2){
  col1 = enquo(col1)
  col2 = enquo(col2)
  data %>% 
    ggvis(col1, col2) %>% layer_points()
}

ggvis_sample(df, b, c)
```
#### ggplot, which uses quoting functions
```{r}
ggplot_sample = function(data, col1, col2){
  col1 = enquo(col1)
  col2 = enquo(col2)
  data %>% 
    ggplot(aes(!! col1, !! col2)) +
  geom_point()
}

ggplot_sample(df, b, c)
```

